Towards Artificial Intelligence for Metaphysics: \\ the Case of the Inconsistency in G\"odel's Ontological Argument

### 2016-01-08 Papers/InProgress/IJCAI/paper.tex
Improve section about embedding of S5

\usepackage{amsmath}
\usepackage{modallogics}
\usepackage{graphicx,url}

\newcommand{\imp}{\rightarrow}
\newcommand{\biimp}{\leftrightarrow}
\newcommand{\allq}{\forall}
\newcommand{\exq}{\exists}
\newcommand{\seq}{\vdash}
\newcommand{\nec}{\Box} % necessarily
\newcommand{\pos}{\Diamond} % possibly

@@

\subsection{Improved Embedding for S5+}

The modal logic S5 requires that the accessibility relation be reflexive, symmetric and transitive. The usual approach to embed S5 would be to use the standard translation for K described above and state these additional conditions as three HOL axioms: 
\begin{itemize}
\item Reflexivity: $\forall x. (r~x~x)$
\item Symmetry: $\forall x. \forall y. (r~x~y) \rightarrow (r~y~x)$ 
\item Transitivity: $\forall x. \forall y, \forall z. (r~x~y) \wedge (r~y~z) \rightarrow (r~x~z)$
\end{itemize}

We consider here a modal logic that we call S5+, which is characterized by the following condition on the accessibility relation:
\begin{itemize}
\item Omni-accessibility: $\forall x. \forall y. (r~x~y)$
\end{itemize}

It is easy to see that S5+ is at least as strong as S5: omni-accessibility entails reflexivity, symmetry and transitivity. S5+, in fact, \emph{strictly stronger} than S5. S5+ only admits complete\footnote{\emph{Complete} in the graph-theoretical sense: for any directed pair of vertices/worlds there is a directed edge connecting them.}} frames, whereas S5 admits non-complete frames as long as all their components are complete.

For S5+, an improved embedding is possible. Omni-accessibility implies that the guarding predicates in the definitions of $\Box$ and $\Diamond$ always hold. Therefore, they can be omitted, and the modal operators can be defined instead merely as:
\begin{itemize}
\item $\Box \varphi \equiv \forall w. (\varphi w)$ 
\item $\Diamond \varphi \equiv \exists w. (\varphi w)$
\end{itemize}

This improved embedding for S5+ is shown in Fig.~\ref{QMLS5}).
\begin{figure}
\centerline{\includegraphics[width=\columnwidth]{./Images/QMLS5.png}}
\caption{Improved Embedding of S5} \label{QMLS5}
\end{figure}

\section{Adequacy of S5+ for Metaphysics}

S5+ is only slightly stronger than S5.  ToDo

\section{Full Automation of Scott's Proof}
Within the modified logic S5 Scott's proof can be fully automated. The
intermediate argumentation steps as not needed anymore to support the
prover, see Fig.~\ref{ScottS5}).
\begin{figure}
\centerline{\includegraphics[width=\columnwidth]{./Images/ScottS5.png}}
\caption{Full Automation of the Ontological Argument in S5 } \label{ScottS5}
\end{figure}

In the modified S5 setting also further results can be automatically verified, which 
failed in previous work: verification of  modal collapse in
Isabelle/HOL with Meson.

>>>>

\usepackage{amsmath}
\usepackage{xspace}
\usepackage{modallogics}
\usepackage{graphicx,url}
\usepackage{txfonts} % needed for \Diamondblack
\usepackage{float}

\newcommand{\imp}{\rightarrow}
\newcommand{\biimp}{\leftrightarrow}
\newcommand{\allq}{\forall}
\newcommand{\exq}{\exists}
\newcommand{\seq}{\vdash}

\newcommand{\Dia}{\Diamond} % possibly
\newcommand{\BlackBox}{\blacksquare}
\newcommand{\BlackDia}{\Diamondblack}

@@

\subsection{Improved Embedding for \SFiveU}

The modal logic \SFive requires that the accessibility relation be reflexive, symmetric and transitive. The usual approach to embed \SFive would be to use the standard translation for K described above and state these additional conditions as three HOL axioms: 
\begin{itemize}
\item Reflexivity: $\forall x. (r~x~x)$
\item Symmetry: $\forall x. \forall y. (r~x~y) \rightarrow (r~y~x)$ 
\item Transitivity: $\forall x. \forall y, \forall z. (r~x~y) \wedge (r~y~z) \rightarrow (r~x~z)$
\end{itemize}

We consider here a modal logic that we call \SFiveU, which is characterized by the following condition on the accessibility relation:
\begin{itemize}
\item Universality: $\forall x. \forall y. (r~x~y)$
\end{itemize}

It is easy to see that \SFiveU is at least as strong as \SFive: universality entails reflexivity, symmetry and transitivity. \SFiveU is, in fact, \emph{strictly stronger} than \SFive. \SFiveU only admits complete\footnote{A graph is \emph{complete} iff there is a directed edge connecting every ordered pair of vertices.} frames, whereas \SFive admits non-complete frames as long as all their components are complete.

For \SFiveU, an improved embedding is possible. Universality implies that the guarding predicates in the definitions of $\Box$ and $\Diamond$ always hold. Therefore, they can be omitted and the accessibility relation can be dispensed altogether. The modal operators can then be defined merely as:
\begin{itemize}
\item $\BlackBox \varphi \equiv \forall w. (\varphi w)$ 
\item $\BlackDia \varphi \equiv \exists w. (\varphi w)$
\end{itemize}

The improved embedding of \SFiveU in Isabelle is shown in Fig.~\ref{QMLS5}.
\begin{figure}
\centerline{\includegraphics[width=\columnwidth]{./Images/QMLS5.png}}
\caption{Improved Embedding of \SFiveU} \label{QMLS5}
\end{figure}

\SFiveU is only slightly stronger than \SFive. Most importantly, $\vDash_{\SFive} \varphi$ iff $\vDash_{\SFiveU} \varphi^{U}$, where $\varphi^U$ is obtained from $\varphi$ by replacing all $\Box$ and $\Dia$ by, respectively, $\BlackBox$ and $\BlackDia$. Therefore, \SFiveU can be considered as adequate for metaphysics as \SFive (cf. \cite{Mattey}, \cite[p. 127]{WilliamsonModalLogicAsMetaphysics}, \cite[ToDo]{DunnHardegree}).

With the improved embedding for \SFiveU, the main theorem T3 can be derived from Scott's version of the axioms fully automatically. The
intermediate argumentation steps as not needed anymore to support the
prover.  This is shown in Fig.~\ref{ScottS5}. There it is also shown that the modal collapse can now be verified in
Isabelle/HOL with Meson. In previous work, with a less efficient embedding for \SFive, this failed.
\begin{figure}
\centerline{\includegraphics[width=\columnwidth]{./Images/ScottS5.png}}
\caption{Full Automation of T3 and MC in \SFiveU } \label{ScottS5}
\end{figure}
\marginpar{ToDo: this picture has to be replaced by one without Isabelle errors}
